---
title: "Extracting Auger yields from PDF manuscript"
author: "Taha Ahmed"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting Auger yields from PDF manuscript}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r packages, echo=FALSE, message=FALSE}
library(knitr)
library(dplyr)
library(magrittr)
library(here)
library(common)
library(reticulate)
library(conflicted)
conflicted::conflict_prefer("filter", "dplyr")
conflicted::conflict_prefer("lag", "dplyr")
```


```{r global_options, echo=FALSE, message=FALSE}
options(
   digits   = 7,
   width    = 84,
   continue = " ",
   prompt   = "> ",
   warn = 0,
   stringsAsFactors = FALSE)
knitr::opts_chunk$set(
   dev        = 'svg',
   fig.width  = 7.10,
   fig.height = 4.39,
   fig.align  = 'center',
   echo       = FALSE,
   eval       = TRUE,
   cache      = FALSE,
   collapse   = TRUE,
   results    = 'hide',
   message    = FALSE,
   warning    = FALSE,
   tidy       = FALSE)
```


So we want to add the Auger yields (which complements the fluorescence yields)
to this dataset from Krause's 1979 paper, which contains a multi-column table
contained on a single page listing the Auger yields for all elements from
B (Z=5) to Ds (Z=110).

It is possible to select the text in the table, but copying and pasting results
in a jumbled mess.

Let's try [the Python way](https://pyvideo.org/pycon-au-2019/extracting-tabular-data-from-pdfs-with-camelot-excalibur.html)
using Camelot.

+ https://camelot-py.readthedocs.io/en/master/user/install-deps.html
+ https://towardsdatascience.com/extracting-tabular-data-from-pdfs-made-easy-with-camelot-80c13967cc88


```{r pysetup, warning=FALSE}
# CAREFUL, if you change pyenv name or path below, make sure to edit ./.Renviron accordingly!
# note, now that we are also sharing some r variables in python, *avoid* using dots in names
pyenv_here <- here::here()
pyenv_name <- "vignette-auger"
pyenv_path <- here::here("man/pyenv", pyenv_name)
reticulate::virtualenv_create(
   # to avoid relying on virtualenv_root(), explicitly provide a path instead of just a name
   envname = pyenv_path,
   # this is just cosmetic, hides repeated PATH warning and makes Python output more readable
   pip_options = c("--no-warn-sscript-location"),
   # note, "camelot-py[base]" (like the docs on pypi says) resulted in only v0.9
   # https://stackoverflow.com/a/67962556/1198249
   packages = c("opencv-python-headless", "ghostscript", "camelot-py"))
# use_python() **doesn't work** in RStudio IDE (claims RETICULATE_PYTHON overrides it,
# despite that env not being set in $HOME or anywhere else on the system), but
# works as expected when run in simple R terminal
# reticulate::use_python(paste0(pyenv_path, "/bin/python"))
# anyway, the solution appears to be to simply set RETICULATE_PYTHON env ourselves
# in .Renviron in this package's root
# py_config() was helpful to understand whether our pyenv was active
# reticulate::py_config()
```


I skimmed through some more docs and the issue queue, and think it might be worth
trying Camelot again with different `flavor` settings.

```{python 'camelot'}
# https://stackoverflow.com/a/61616750
# import camelot fails in Python terminal (inside pyenv, on luxor), with:
# import-im6.q16: unable to open X server `' @ error/import.c/ImportImageCommand/346.
# headless problem? yes, fixed by installing opencv-python-headless and importing camelot.io
# https://stackoverflow.com/a/67962556
import camelot.io as camelot
import cv2
import os
# print(r.pyenv_here)
# https://stackoverflow.com/questions/7132861/build-the-full-path-filename-in-python
pdf_path = os.path.join(r.pyenv_here, "inst/extdata/Krause1979-p13.pdf")
#print(pdf_path)
# camelot.read_pdf() failed: ModuleNotFoundError: No module named 'ghostscript'
# OK, resolved by installing pip package ghostscript
auger_table = camelot.read_pdf(pdf_path, flavor='stream')
# auger_table
# read_pdf(..., flavor='lattice') -> empty list <TableList n=0>
# read_pdf(..., flavor='stream') -> <TableList n=1>
auger_table[0].parsing_report
csv_path = os.path.join(r.pyenv_here, "inst/extdata/Krause1979-p13.csv")
auger_table[0].to_csv(csv_path)
```

Not too bad by Camelot. There is definitely a bunch of transcription errors,
but the overall layout of the tabular is preserved in the CSV.
Will require a lot of manual cleanup, but still better than transcribing it
manually from scratch.
Although I wonder if another approach like Tesseract OCR would give better
transcription results?

(Note, since I already invested time and effort here in getting python venv working
inside this R package vignette, I'd prefer a Python-based alternative to Camelot).


+ https://towardsdatascience.com/extracting-tabular-data-from-pdfs-made-easy-with-camelot-80c13967cc88
+ https://pypi.org/project/camelot-py
+ https://github.com/camelot-dev/camelot
+ https://cran.r-project.org/web/packages/reticulate/vignettes/python_packages.html
+ https://www.lukaskawerau.com/rmarkdown-with-python-and-virtual-envs/
+ https://rstudio.github.io/reticulate/articles/r_markdown.html
+ https://github.com/JaidedAI/EasyOCR
+ https://github.com/jbarlow83/OCRmyPDF
+ https://pypi.org/project/pytesseract/
+ https://tesseract-ocr.github.io/tessdoc/Home.html
